A sheaf--function dictionary for algebraic tori over local fields

### 2013-08-20 shanghai.tex
moving computers ....

The proof of Proposition~\ref{prop:functorialG} requires the results from the next section.

@@

\David{insert some commentary on this proposition: it will be very useful to us}

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gqcs{L},\mu,\phi) \mapsto (\bf^*\gqcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \QC(G) \to \QC(H)$ such that
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \QCiso{G} \rar{f^*} \arrow[swap]{d}{\trFrob{G}} & \QCiso{H} \dar{\trFrob{H}} \\
   G(\Fq)^* \rar & H(\Fq)^*
  %\Hom_\text{grp}(G(\Fq),\EEx) \rar & \Hom_\text{grp}(H(\Fq),\EEx)
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

We will also need the Greenberg functor, 
introduced in \cite{greenberg:61} and \cite{greenberg:63a}, 
as a functor from the category of schemes locally of finite type 
over an Artinian local ring with perfect residue field $k$ 
to the category of schemes locally of finite type over $k$. 
This functor respects open immersions, closed immersions, 
etale morphisms, smooth morphisms and geometric components.
Besides Greenberg's original papers, 
see \cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6} 
and the first few sections of \cite{stasinski:12a} 
for background on the Greenberg functor.

@@

\begin{proof}
Weil restriction behaves properly with respect to formation of Néron models:
there is a canonical isomorphism
\[
\underline{\Res_{K'/K}(T')} \iso \Res_{\mathcal{O}_{K'}/\OK}(\underline{T'}).
\]
And Weil restriction behaves properly with respect to the Greenberg transform:
there is a canonical isomorphism 
\Clifton{References needed!}
\[
\Gr_{n} \Res_{\mathcal{O}_{K'}/\OK}(\underline{T'})
\iso
\Res_{k'/k} \Gr_n(\underline{T'}),
\]
where $k'$ is the residue field for $K'$.
And the Greenberg tranforms respects products:
\[
\Gr_n(\underline{T'}) = \Gr_n(\underline{T})\times_{k} {k'}.
\]
Now use Proposition~\ref{prop:csbe}.
\end{proof}

>>>>

The proof of Proposition~\ref{prop:functorialG} will be given at the end of the next section.

@@

%\David{insert some commentary on this proposition: it will be very useful to us}
As a step toward the proof of Proposition~\ref{prop:functorialG} we show that every morphism  $H \to G$ of smooth group schemes locally of finite type over $\Fq$ induces a monoidal functor from quasicharacter sheaves on $G$ to quasicharacters on $H$. This simple but important result will also play a role in the proof of many other propositions in this paper.

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gqcs{L},\mu,\phi) \mapsto (\bf^*\gqcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \QC(G) \to \QC(H)$ such that
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \QCiso{G} \rar{f^*} \arrow[swap]{d}{\trFrob{G}} & \QCiso{H} \dar{\trFrob{H}} \\
   G(\Fq)^* \rar & H(\Fq)^*
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

We will also need the Greenberg functor, 
introduced in \cite{greenberg:61} and \cite{greenberg:63a}, 
as a functor from the category of schemes locally of finite type 
over an Artinian local ring with perfect residue field $k$ 
to the category of schemes locally of finite type over $k$. 
This functor respects open immersions, closed immersions, 
etale morphisms, smooth morphisms and geometric components.
Besides Greenberg's original papers, see [DG70, V, § 4, no. 1])
see \cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch. 9, \S 6} 
and the first few sections of \cite{stasinski:12a} 
for background on the Greenberg functor.

@@

\begin{proof}
Weil restriction behaves properly with respect to formation of Néron models:
there is a canonical isomorphism
\[
\underline{\Res_{K'/K}(T')} \iso \Res_{\mathcal{O}_{K'}/\OK}(\underline{T'}).
\]
And Weil restriction behaves properly with respect to the Greenberg transform:
there is a canonical isomorphism 
\Clifton{References needed!}
\[
\Gr_{n}\left( \Res_{\mathcal{O}_{K'}/\OK}(\underline{T'})\right)
\iso
\Res_{k'/k} \Gr_n(\underline{T'}),
\]
where $k'$ is the residue field for $K'$.
And the Greenberg tranforms respects products:
\[
\Gr_n(\underline{T'}) = \Gr_n(\underline{T}\times_{k} {k'}) \iso \Gr_n(\underline{T})\times_{k} {k'} = \Gr_n(\underline{T})_{k'}.
\]
Thus, the 
Now use Proposition~\ref{prop:csbe}.
\[
\begin{tikzcd}
\QCiso{\Res_{k'/k}(G_{k'})} \arrow[two heads]{r}{\iota^*} \dar[swap]{\trFrob{\Res_{k'/k}(G_{k'})}} & \QCiso{G} \dar{\trFrob{G}} \\
G(k')^* \arrow[two heads]{r}{i_{k'/k}^*} & G(k)^*.
\end{tikzcd}
\]

\end{proof}
